### 2024-11-27 paper/paper.qmd
Updating introduction and model sections

---
title: "Valuing Real Estate in Seattle Using Property Characteristics: A Machine Learning Approach"
subtitle: "Subtitle"
author:
  - Sameeck Bhatia
date: today
date-format: long
thanks: "Code and data are available at [https://github.com/SameeckBhatia/Seattle-Real-Estate](https://github.com/SameeckBhatia/Seattle-Real-Estate)"
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi faucibus ipsum dolor, in tincidunt justo tempor sit amet proin non turpis a orci imperdiet pretium. Aenean aliquam eros eu consectetur cursus. Proin vel tempor ante vestibulum felis nisi, ultricies sed porttitor in, viverra nec augue. Sed convallis a tortor in pretium. Vivamus metus ipsum, faucibus vitae turpis sed, sodales auctor dolor sed a orci id orci pulvinar tincidunt sed ut tellus. Morbi venenatis tempor arcu, a auctor dolor congue sit amet. Nam auctor tincidunt ligula rutrum aliquam. Sed facilisis at nulla et posuere."
toc: true
number-sections: true
bibliography: references.bib
format: 
  pdf:
    fig.width: 6
    fig.height: 4.5
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| include: FALSE

library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(arrow)
library(stargazer)
library(ggpubr)
library(modelsummary)
```

# Introduction

Seattle’s real estate market has experienced significant fluctuations since the start of the millennium. Despite these booms and busts, property prices in Seattle remain among the highest in the United States. Demand has steadily increased, and even during periods of lower demand, the market continues to attract buyers. At current price levels, more potential homebuyers are seeking reliable information to find the best deals. With advancements in technology and the increasing availability of data, tools for assessing property values have become essential. One critical tool is an estimate of a property’s fair value, which can help buyers determine whether they are overpaying or underpaying relative to the market.

This paper focuses on developing an accurate and accessible property valuation model tailored to Seattle’s housing market. The model incorporates current market data to provide fair value estimates for individual properties. The goal is to inform readers about the key factors driving real estate prices in Seattle while empowering them with a tool to make better home-buying decisions. By applying machine learning and statistical techniques, the model aims to ensure accuracy and reflect the intrinsic value of properties rather than merely their market prices.

Existing valuation tools, such as Zillow’s “Zestimate” and Redfin’s proprietary estimates, have limitations. These models are often closed-source, leaving potential users unable to access the methodology or even the results without cost. This creates a significant gap in the availability of open, free, and transparent valuation tools. The absence of such tools denies buyers the advantage of comprehensive, unbiased information to assess property values independently. To address this gap, data was collected from Redfin’s semi-public dataset on current Seattle property listings. The analysis was conducted using the R programming language for data cleaning, testing, model creation, and result interpretation.

**TO-DO: ** Preliminary findings, significance, estimand

The paper is structured as follows: @sec-data introduces the dataset and variables. @sec-model outlines the model design and its significance. @sec-results presents the model’s predictions and actionable insights for buyers. Finally, @sec-discuss explores potential applications of the model in other cities, factors influencing property prices, and the generalizability of the model.

# Data {#sec-data}

## Overview

The data used in this paper comes from Redfin, a real estate brokerage and mortgage company, and represents current property prices across Seattle. While the data was collected from @redfin, similar data could have been sourced from competitors like Zillow and Realtor.com, with Zillow being the largest of the three. If data from Zillow or Realtor.com had been used, the listings might vary slightly since each platform likely features different sellers and listings. However, Redfin was chosen as the source because it is the only brokerage firm that allows public downloads of its listings, helping to avoid violations of data extraction policies.

The dataset includes only properties within Seattle’s official city boundaries and covers condos, townhouses, and single-family homes currently on the market. The raw data contains 27 variables, 18 of which have been included in the cleaned dataset. These include variables such as MLS number, number of bedrooms, neighborhood, and geographic coordinates. Additionally, some variables were constructed specifically for this analysis, such as `half_bath`, `property_age`, and `price_sqft`. The first two were added to enhance the dataset and valuation model, while `price_sqft` was created as an alternate response variable. Further details about these variables can be found in Appendix A. The data was analyzed using `R` [@R] and the `tidyverse` [@tidyverse] package, while visualizations have been created using `tidyverse`.

## Summary Statistics

```{r}
#| output: asis

data <- read_parquet("../data/cleaned_data/cleaned_data.parquet") |>
  as.data.frame()

data |>
  select(!c("zip_code", "latitude", "longitude")) |>
  stargazer(type = "latex", header = FALSE, digits = 2, median = TRUE, 
            label = "tbl-stats")
```

Table \ref{tbl-stats} presents summary statistics for all original and derived numeric variables in the dataset. The mean number of bedrooms is 2.69, with a median of 3, while the mean number of bathrooms is 2.03, with a median of 2. This indicates that the typical listing has around three bedrooms and two bathrooms, commonly seen in townhouses and single-family homes. The average property size is approximately 1,710 square feet, with a median of 1,373 square feet, suggesting a positive skew in property size due to a few larger homes in the dataset. The average time a property remains on the market is around 64 days (just over 2 months), with a maximum of 878 days (nearly 2.5 years), indicating low demand in Seattle’s real estate market, especially since the data contains only active listings. The mean and median property prices are \$1,107,431 and \$775,000, respectively, while the mean and median price per square foot are \$631.41 and \$590.71, respectively. These figures highlight Seattle as one of the most expensive residential markets in the United States.

## Measurement

In the United States, buyers and sellers have the freedom to select the real estate agent or brokerage firm they wish to work with for transactions. Consequently, agents often represent multiple listings within their region or city. Agents receive detailed information on each property from real estate appraisers, who measure variables such as the number of bedrooms, bathrooms, and square footage [@nar]. These measurement practices, except for price, are strictly regulated to ensure accuracy for all stakeholders.

For property prices, appraisers typically estimate values based on the prices of recently sold comparable properties and the specific characteristics of the property. This valuation process is less regulated, as it serves primarily as a reference point for buyers and sellers. Real estate agents may gather price estimates from multiple appraisers to calculate an average. The prices observed in the data, although guided by these values, are ultimately set by the seller. All this information is uploaded to the Multiple Listing Service (MLS), a private database accessible only to agents and brokerage firms via subscription fees [@bankrate]. However, U.S. laws allow companies like Redfin, Trulia, and Zillow to extract and share MLS data with the public, fostering competition and transparency.

## Outcome Variables

The primary goal of this paper is to estimate the outcome variable, price, using a valuation model. This variable represents the market value of active listings at the time of data collection, focusing exclusively on properties located in Seattle. @fig-out provides insights into the distribution of property prices. Plot A shows that property prices are highly skewed, with a maximum value near \$40,000,000 and several listings exceeding \$5,000,000. To better understand the distribution, Plot B presents the logarithmic transformation of property prices. While the transformed distribution is less extreme, it remains skewed, indicating the presence of properties with exceptionally high valuations in the Seattle real estate market.

```{r fig-out, fig.width=7, fig.height=2.5, fig.cap='Heavily Skewed Property Prices and Skewed Still for Logarithmic Prices'}
p1 <- data |>
  select(price) |>
  ggplot(aes(x = price)) +
  geom_histogram(binwidth = 1000000, fill = "beige", color = "black") +
  labs(x = "Price ($)", y = "Count")

p2 <- data |>
  select(price) |>
  ggplot(aes(x = log(price))) +
  geom_histogram(binwidth = 0.25, fill = "beige", color = "black") +
  labs(x = " ln(Price)", y = "Count")

ggarrange(p1, p2, labels = c("A", "B"))
```

## Predictor Variables

**Property Type (`property_type`):** This categorical variable identifies the type of property for each observation, based on classifications set by the MLS. The three main property types included in the analysis are "Condo/Co-op," "Single Family," and "Townhouse." These classifications are critical for valuation, as different property types possess distinct features that influence their market value.

**Number of Bedrooms (`beds`):** This numeric variable represents the count of full bedrooms in a property, as measured by appraisers. It is a key factor in the analysis, as more bedrooms often correlate with greater living space and the potential to accommodate larger households.

**Property Size (`sqft`):** Measured in square feet, this variable reflects the total size of the property and is determined by appraisers. As the United States predominantly uses the imperial system, square footage is a standard unit. Larger properties generally hold more value, making this a significant predictor in property valuation.

**Property Age (`property_age`):** This derived variable calculates the age of a property in years, based on the difference between the year of data collection and the `year_built`. Age is a vital consideration in determining price, as newer properties are often smaller due to rising construction costs and increasingly strict zoning regulations.

@fig-pred1 represents the distributions of property types and number of bedrooms. Plot C shows that single-family homes are the most common, followed closely by condos, with townhouses being the least prevalent. This likely reflects the mix of apartments and houses present in Seattle's real estate market. Plot D highlights that properties with 2 or 3 bedrooms are most frequent, while some properties even have north of 7 bedrooms. This is likely because condos typically feature 1-2 bedrooms, whereas single-family homes and townhouses commonly have 3-4 bedrooms.

```{r fig-pred1, fig.width=7, fig.height=2.5, fig.cap='Single Family Homes and 2 or 3 Bedrooms Properties Most Listing on the Market'}
p3 <- data |> 
  ggplot(aes(x = property_type)) +
  geom_bar(fill = "beige", color = "black") +
  labs(x = "Property Type", y = "Count")

p4 <- data |> 
  mutate(beds = as.factor(beds)) |>
  ggplot(aes(x = beds)) +
  geom_bar(fill = "beige", color = "black") +
  labs(x = "Beds", y = "Count")

ggarrange(p3, p4, labels = c("C", "D"))
```

@fig-pred2 displays the distributions of property size and age. Plot E shows that property size is positively skewed, with most properties under 2,500 square feet and a peak between 700 and 1,200 square feet. Properties exceeding 5,000 square feet likely represent luxury homes or mansions. Plot F indicates that property age is also positively skewed, with the most common properties being less than 10 years old. The second most common group is 20-30 years old, likely reflecting construction surges before the 2008 financial crisis and real estate market downturn.

```{r fig-pred2, fig.width=7, fig.height=2.5, fig.cap='Skewed Distribution of Sizes and Most Properties Built Over the Last Decade'}
p5 <- data |>
  ggplot(aes(x = sqft)) + 
  geom_histogram(binwidth = 500, fill = "beige", color = "black") +
  labs(x = "Size (sqft)", y = NULL)

p6 <- data |> 
  ggplot(aes(x = property_age)) + 
  geom_histogram(binwidth = 10, fill = "beige", color = "black") +
  labs(x = "Property Age (years)", y = NULL)

ggarrange(p5, p6, labels = c("E", "F"))
```

\newpage

# Model {#sec-model}

## Overview

The model uses a multiple linear regression approach to estimate property values in Seattle. It incorporates features like the number of bedrooms (`beds`), square footage (`sqft`), homeowner association fees (`hoa_month`), and an interaction term (`sqft` $\times$ `beds`) to predict prices. These characteristics were selected for their significance in influencing property valuations. The final model was chosen after thorough validation, achieving a good balance of accuracy and interpretability.

## Setup

The linear regression model can be represented as:

$$
\text{price} = \beta_0 + \beta_1 \cdot \text{beds} + \beta_2 \cdot \text{sqft} + \beta_3 \cdot \text{hoa\_month} + \beta_4 \cdot (\text{sqft} \cdot \text{beds})
$$

, where the coefficients are described as:

**$\beta_0$: Intercept**

The predicted value of `price` when all predictors ($\text{beds}$, $\text{sqft}$, $\text{hoa\_month}$, $\text{sqft} \cdot \text{beds}$) are zero.

**$\beta_1$: Coefficient for `beds`**

The effect of adding one more bedroom ($\text{beds}$) on `price`, holding other factors constant (except for the interaction term).

**$\beta_2$: Coefficient for `sqft`**

The effect of increasing square footage ($\text{sqft}$) by one unit on `price`, holding other factors constant (except for the interaction term).

**$\beta_3$: Coefficient for `hoa_month`**

The effect of a one-unit increase in monthly homeowner's association fees (`hoa_month`) on `price`, holding other factors constant.

**$\beta_4$: Interaction between `sqft` and `beds`**

Represents how the relationship between $\text{sqft}$ and `price` changes depending on the number of bedrooms ($\text{beds}$).

## Limitations

This model has a few limitations that should be considered when interpreting its predictions. First, it is trained on cross-sectional data rather than longitudinal data. As a result, the model is designed to provide accurate valuations for properties in the near future (typically less than a year), assuming minimal price fluctuations. Significant market changes over time would reduce the model's accuracy, requiring frequent updates with new cross-sectional data to maintain reliability. Additionally, the model is specifically tailored to Seattle’s real estate market. While it might perform adequately in nearby cities with similar market characteristics, it is unlikely to generalize well to cities in other states, such as Los Angeles or New York, due to differing property attributes and pricing dynamics in those regions.

## Justification

The four features included in the model are `beds`, `sqft`, `hoa_month`, and `sqft:beds`, as they were the most significant in influencing the model's property valuations. The beds feature was selected because the number of bedrooms is a key determinant of a property's utility and appeal to buyers, directly impacting its market value. The sqft feature, representing the total square footage, is a fundamental metric for assessing a property's size and, consequently, its worth. The hoa_month feature accounts for monthly homeowner association fees, which can significantly affect the affordability and desirability of properties, particularly in condominiums or communities with shared amenities. Finally, the interaction term sqft:beds captures the relationship between the size of the property and the number of bedrooms, highlighting how the distribution of space impacts valuation.

```{r}
full_model <- readRDS("../models/full_model.rds")
reduced_model <- readRDS("../models/reduced_model.rds")
final_model <- readRDS("../models/final_model.rds")
```

```{r #tbl-summ}
modelsummary(list(
  # "Full" = full_model,
  # "Reduced" = reduced_model,
  "Final Model" = final_model), output = "kableExtra",
  gof_omit = "BIC|AIC") |>
  column_spec(2, width = "2.5in")
```

## Interpretation

@tbl-summ shows the values of the model's coefficients mentioned in the setup. The intercept of -341,047.42 suggests that, without any additional predictors, property prices would begin at this negative baseline, though this value has limited standalone interpretation. The coefficient for beds is -159,574.45, meaning that, holding other factors constant, each additional bedroom is associated with a significant reduction in price. However, this effect is modified by the interaction term beds $\times$ sqft, which has a negative coefficient of -32.64, indicating that larger properties with more bedrooms have diminishing price returns. The sqft variable has a strong positive impact (1152.06 per additional square foot), highlighting size as a key driver of value. The hoa_month coefficient (267.83) shows a modest positive association with price, possibly reflecting higher costs in premium communities. With an $R^2$ of 0.654, the model explains 65.4\% of the variance in property prices, and an RMSE of 719,452.71, the summary suggests a decent fit.

## Validation

The linear model was created using base R [@R] to fit the data and generate predictions, while the MLMetrics package [@mlmetrics] was used to evaluate performance. A train-test split of 70:30 was created with the rsample package [@rsample], with the model trained on the larger portion of the data and validated using out-of-sample testing on the remaining data. Key evaluation metrics included the Root Mean Square Error (RMSE) and the $R^2$ score. The model achieved an RMSE of 1,850,720, indicating the average prediction error in dollar terms, and an $R^2$ score of approximately 0.281, reflecting the proportion of variance in property prices explained by the model. Detailed diagnostics and further information are provided in the appendix.

## Alternate Models

Several alternative models were considered before selecting the final one. The first, a “full model,” included all variables in the cleaned dataset, achieving an RMSE of 1,864,966 and an $R^2$ score of 0.269. While this model used all available information, many variables were statistically insignificant. A reduced model was then tested by retaining only significant variables, resulting in an RMSE of 1,865,188 and an $R^2$ score of 0.269. Though concise, it performed similarly to the full model. The final model, incorporating an interaction term, outperformed both alternatives with an RMSE of 1,850,720 and an $R^2$ score of 0.281.

# Results {#sec-results}

## Result 1

```{r, fig.align='center'}
palmerpenguins::penguins |>
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
  geom_point()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse accumsan elementum diam sed ullamcorper. Integer sodales posuere mauris, sit amet blandit tortor scelerisque non. Nulla ut sapien vel orci scelerisque varius malesuada vel lorem. Integer sit amet enim mattis nisi porttitor sodales nec vitae odio. Vivamus eget rutrum erat. Ut dictum quis ligula non euismod. Fusce sit amet egestas nibh. In rutrum tortor sed nulla volutpat facilisis. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc non sapien lacus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Fusce nec magna nec leo convallis dignissim ut porta justo. Curabitur nec faucibus orci. Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum. Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas. Morbi eget odio nisi. Curabitur posuere scelerisque libero, non pharetra ex scelerisque.

## Result 2

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = sex, y = body_mass_g, fill = sex)) +
  geom_boxplot(width = 0.5) +
  coord_flip()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse accumsan elementum diam sed ullamcorper. Integer sodales posuere mauris, sit amet blandit tortor scelerisque non. Nulla ut sapien vel orci scelerisque varius malesuada vel lorem. Integer sit amet enim mattis nisi porttitor sodales nec vitae odio. Vivamus eget rutrum erat. Ut dictum quis ligula non euismod. Fusce sit amet egestas nibh. In rutrum tortor sed nulla volutpat facilisis. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc non sapien lacus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Fusce nec magna nec leo convallis dignissim ut porta justo. Curabitur nec faucibus orci. Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum. Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas. Morbi eget odio nisi. Curabitur posuere scelerisque libero, non pharetra ex scelerisque.

## Result 3

```{r, fig.align='center'}
palmerpenguins::penguins |>
  drop_na() |>
  ggplot(aes(x = species, y = flipper_length_mm, fill = species)) +
  geom_violin(alpha = 0.5) +
  geom_boxplot(width = 0.25) +
  coord_flip()
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Suspendisse accumsan elementum diam sed ullamcorper. Integer sodales posuere mauris, sit amet blandit tortor scelerisque non. Nulla ut sapien vel orci scelerisque varius malesuada vel lorem. Integer sit amet enim mattis nisi porttitor sodales nec vitae odio. Vivamus eget rutrum erat. Ut dictum quis ligula non euismod. Fusce sit amet egestas nibh. In rutrum tortor sed nulla volutpat facilisis. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nunc non sapien lacus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Fusce nec magna nec leo convallis dignissim ut porta justo. Curabitur nec faucibus orci. Sed fringilla elit vel sapien rutrum, ut egestas nisl bibendum. Donec faucibus suscipit enim, porttitor suscipit ex commodo egestas. Morbi eget odio nisi. Curabitur posuere scelerisque libero, non pharetra ex scelerisque.

\newpage

# Discussion {#sec-discuss}

## Discussion 1
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

## Discussion 2

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

## Discussion 3

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Sed eget purus eleifend, vehicula dolor vitae, aliquet nibh. Curabitur ac leo non nunc ultrices dignissim. Phasellus faucibus elit id nulla pretium volutpat. Proin ullamcorper, nulla sit amet tempor faucibus, mi ligula interdum urna, eu condimentum arcu augue nec leo. Curabitur vulputate imperdiet nulla eu finibus. Vivamus sem lectus, commodo in lectus quis, posuere laoreet quam. Etiam nunc libero, faucibus nec porta a, dignissim a nisi. Phasellus laoreet elementum turpis a vestibulum. Quisque in urna varius, lobortis eros eu, blandit sapien. In lobortis orci id nunc facilisis, non tempor nisl ultricies. Sed pellentesque tortor cursus ipsum fringilla interdum. Cras lacinia tempor placerat. Proin eu elit elit. Sed molestie dolor et posuere sodales.

Fusce aliquam tempus tincidunt. Nulla viverra non nibh consectetur blandit. Proin malesuada, nunc et accumsan pellentesque, odio felis molestie augue, nec gravida lectus ipsum ut lectus. Donec at malesuada purus. Cras ultrices vitae metus eget aliquet. Aliquam sollicitudin elit placerat, porta mauris in, semper felis. Nulla vestibulum dolor sed tempor efficitur. In bibendum libero ex, in sollicitudin arcu maximus porttitor. Pellentesque ac sapien sit amet magna ornare auctor. In hac habitasse platea dictumst. Cras ut urna dapibus elit mattis aliquam non sed augue. Suspendisse lacinia ornare consectetur. Quisque ultricies tortor non ullamcorper porta. Mauris ac leo quis lacus tincidunt cursus ut id felis.

\newpage

\appendix

# Appendix {-}

# Additional Data Details

```{r}
data_dictionary <- data.frame(
  Variable = c(
    "SALE TYPE", "SOLD DATE", "PROPERTY TYPE", "ADDRESS", "CITY", 
    "STATE OR PROVINCE", "ZIP OR POSTAL CODE", "PRICE", "BEDS", 
    "BATHS", "LOCATION", "SQUARE FEET", "LOT SIZE", "YEAR BUILT", 
    "DAYS ON MARKET", "$/SQUARE FEET", "HOA/MONTH", "STATUS", 
    "NEXT OPEN HOUSE START TIME", "NEXT OPEN HOUSE END TIME", 
    "URL", "SOURCE", "MLS#", "FAVORITE", "INTERESTED", "LATITUDE", "LONGITUDE"
  ),
  Description = c(
    "Type of sale (e.g., MLS listing, new construction home, new construction plan).",
    "Date when the property was sold.",
    "Type of property (e.g., condo, single-family house, townhouse).",
    "Full address of the property.",
    "City where the property is located.",
    "State or province where the property is located.",
    "ZIP code of the property location.",
    "Sale price of the property.",
    "Number of bedrooms in the property.",
    "Number of bathrooms in the property.",
    "Neighbourhood of the property.",
    "Total square footage of the property.",
    "Lot size in square feet",
    "Year the property was built.",
    "Number of days the property has been listed on the market.",
    "Price per square foot of the property.",
    "Monthly Homeowners Association (HOA) fee, if applicable.",
    "Current status of the property (e.g., sold, pending, active).",
    "Start time of the next scheduled open house, if available.",
    "End time of the next scheduled open house, if available.",
    "URL to additional property details.",
    "Source of the property data (e.g., MLS, Zillow).",
    "Multiple Listing Service (MLS) identification number for the property.",
    "Indicates whether the property is marked as a favorite (e.g., Y/N).",
    "Indicates whether the user has expressed interest in the property (e.g., Y/N).",
    "Latitude of the property location for geospatial analysis.",
    "Longitude of the property location for geospatial analysis."
  )
)

kable(data_dictionary, col.names = c("Variable", "Description"), align = "l")
```

# Additional Model Details

```{r, fig.width=6, fig.height=3}
par(mfrow = c(1, 2))

plot(fitted(final_model), resid(final_model), main = "Residual-Fitted Plot",
     xlab="Fitted", ylab = "Residual")
qqnorm(resid(final_model));qqline(resid(final_model))
```

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis vitae pretium velit. Etiam fermentum enim ac venenatis fringilla. Nulla imperdiet metus ac ligula ultricies, at cursus eros varius. Curabitur commodo pretium ornare. Pellentesque ultrices sapien ut lectus viverra placerat. Mauris nec interdum tortor. Cras posuere turpis sed mi blandit dictum. Etiam accumsan ipsum quis sem ornare cursus. Aenean ligula mauris, vulputate a volutpat dapibus, consectetur a lacus. Nullam ornare placerat risus, in luctus urna. Ut molestie dignissim turpis, eget tincidunt arcu. Vivamus auctor fermentum lectus, eu vulputate tortor volutpat et. Aenean aliquam nulla eget felis faucibus faucibus ut nec nisi.

```{r}
modelsummary(list(
  "Full" = full_model,
  "Reduced" = reduced_model), 
  output = "kableExtra") |>
  column_spec(2, width = "2in")
```

\newpage

# References